---
title: "Thermoregulation Analysis, Streamlined"
author: "Callin Switzer"
date: "15 Jan 2018"
output:
  html_document: default
  html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# setup
```{r, warning = FALSE, message=FALSE}
# Callin Switzer
# Looking over James' code for analysis of thermoregulation data

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
packages <- c("ggplot2", "plyr", "viridis", "lme4", "tidyr", 
              "dplyr", "lubridate", "signal", "zoo", "effects", 
              "gamm4", "MuMIn", "stargazer", "broom", "plyr")
ipak(packages)


# set ggplot theme
theme_set(theme_classic())

# set wd
setwd("/Users/cswitzer/Documents/GitRepos/thermoRegImidacloprid")

# make a figures directory, if it doesn't exist
dir.create(file.path(getwd(), "figuresStreamlined"), showWarnings = FALSE)
figDir <- file.path(getwd(), "figuresStreamlined")
dataDir <- '/Users/cswitzer/Documents/GitRepos/thermoRegImidacloprid/data'

# print session info
print(paste("Last update:", Sys.time(), Sys.timezone()))
print(sessionInfo())
```



# Load and manipulate data
```{R}
# Define temperature set point
sp <- 32.5 

#___________________________________________
# Load data from each cohort
#___________________________________________
# cohort 1
data <- read.csv(file.path(dataDir, 'summaryDataC1.csv'))
treatList1 <- c('c', 't', 'c', 't', 't', 'c')
data$cohort = 1


# cohort #2
data2 <-  read.csv(file.path(dataDir, 'summaryDataC2.csv'))
treatList2 <- c('c', 't', 't', 't', 'c', 'c')
data2$cohort = 2

# cohort #3
data3 <-  read.csv(file.path(dataDir, 'summaryDataC3.csv'))
treatList3 <- c('t', 'c', 'c', 'c', 't', 't')
data3$cohort = 3

#___________________________________________
# combine cohorts
#___________________________________________
combData <- rbind(data, data2, data3)

# converts time to a proportion of a day (0 is midnight, 0.5 is noon)
combData$time1 <- (combData$time) %% 1

# add day
combData$day = floor(combData$time)

# Make the "time" variable into a datetime format -- 
# dates are all set 1970-01-01
# 86400 = 24 * 60 * 60, which converts days to seconds
combData$time2 <- format(as.POSIXct((combData$time1) * 86400, 
                                    origin = "1970-01-01", tz = "UTC"), 
                         "%H:%M:%S")

# time3 is the hour as a number, for example 14.5 is equal to 14:30
combData$time3 <- as.numeric(substr(combData$time2, 1,2)) +
  as.numeric(substr(combData$time2, 4,5))/60 + 
  as.numeric(substr(combData$time2, 7,8)) / 60 / 60

#___________________________________________________________
# smooth ambient temperature and calculate increasing vs.
# decreasing
#___________________________________________________________

dataIncDec <- lapply(unique(combData$cohort), function(ii){
  foo <- combData[combData$cohort == ii, ]
  
  plot(foo$ambient, x=foo$time, pch = ".")
  foo$rollMeanAmbient <- rmn <- rollapply(foo$ambient, width = 482, FUN = mean, align = "center", partial = TRUE)
  lines(y = rmn, x = foo$time, col = 'orange', lwd = 1)
  
  # what is 482 points?  -- answer: 3 hours
  print (mean(sapply(1:100, function(jj){
    foo$time3[jj + 482] - foo$time3[jj]})))
  
  foo$tempDer <- c(NA, diff(rmn)) # add NA to the beginning, since diff reduces size by 1
  foo$tempIncrease <- ifelse(foo$tempDer >= 0, yes = "Increasing", no = "Decreasing")
  
  foo$tempIncrease_noSM <- c(NA, diff(foo$ambient))
  foo$tempIncrease_noSmooth <- ifelse(foo$tempIncrease_noSM >= 0, yes = "Increasing", no = "Decreasing")
  
  return(foo)
})

# merge back into dataframe
df3 <- rbind.fill(dataIncDec)
combData <- merge(combData, df3, all.x = TRUE)
```

# Visualization
```{R, cache = TRUE}
# visualize increasing vs. decreasing
# downsample so that output image isn't too big to put into Adobe or MS word


p2 <- ggplot(drop_na(combData[seq(1, nrow(combData), length.out = 20000), ]), aes(x = time, y = ambient)) + 
  geom_rect(aes(xmin = time, xmax = time, ymin = 0, ymax = 1, color = tempIncrease, fill = tempIncrease), size = 0.1) +
  geom_line(size = 0.1) + 
  geom_line(aes(y = rollMeanAmbient, color = tempIncrease, group = 1), alpha = 0.5) +
  facet_grid(cohort~., labeller = labeller(.rows = label_both, .cols = label_both)) + 
  scale_color_viridis(discrete = TRUE,begin = 0.4, end = 1, name = "Temperature") +
  scale_fill_viridis(discrete = TRUE,begin = 0.4, end = 1, name = "Temperature") +
  theme(legend.position = "right") + 
  labs(x = "Time (days)", y = "Ambient temperature (°C)")
p2

png(file.path(figDir, "ambientIncreasingDecreasing.png"), width = 10, height= 7, units = "in", res = 250)
p2 # figure s6
dev.off()

pdf(file.path(figDir,'ambientIncreasingDecreasing.pdf'), width = 10, height= 7)
p2
dev.off()
```


# Convert data to long format
```{R}
#___________________________________________
# convert combined dataset to long format
# and add some extra variables
#___________________________________________
data_long <- gather(combData, condition, temp, c1brood:c6air, factor_key=TRUE)


# the colony variable includes cohort_colony information
data_long$colony = paste(data_long$cohort,  
                         substr(data_long$condition,start = 2,2 ), 
                         sep = "_")

# location is either air or brood
data_long$location = substr(data_long$condition,start = 3,1000 )

# dayTime is either day or night
data_long$dayTime = ifelse(data_long$time3> 6 & data_long$time3 < 20, "day", "night")

# insert treatment
data_long$treatment <- mapvalues(data_long$colony, 
                                 from = c('1_1' , '1_2' , '1_3' , '1_4' , '1_5' , 
                                          '1_6' , '2_1' , '2_2' , '2_3' , '2_4' , 
                                          '2_5' , '2_6' , '3_1' , '3_2' , '3_3' , 
                                          '3_4' , '3_5' , '3_6'), 
                                 to = c(treatList1, treatList2, treatList3))

data_long$treatment <- mapvalues(data_long$treatment, from = c("c", "t"), 
                                 to = c("control_grp", "treatment_grp"))


#drop NA's -- 36 is from 18 colonies and air/brood measurements in each colony
sum(is.na(data_long$tempIncrease)) 
data_long <- data_long[!is.na(data_long$tempIncrease), ]
```


# Separate datasets into air and brood

```{R}
# new dataset that includes only brood temp (not air)
brooddta <- data_long[data_long$location == "brood", ]
airdta <- data_long[data_long$location == "air", ]
```


# Estimate VIF
```{R}
vif.mer <- function (fit) {
  ## adapted from rms::vif
  
  v <- vcov(fit)
  nam <- names(fixef(fit))
  
  ## exclude intercepts
  ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
  if (ns > 0) {
    v <- v[-(1:ns), -(1:ns), drop = FALSE]
    nam <- nam[-(1:ns)]
  }
  
  d <- diag(v)^0.5
  v <- diag(solve(v/(d %o% d)))
  names(v) <- nam
  v
}

set.seed(17444)
brooddta_sm <- sample_n(brooddta, 10000, replace = FALSE)
brooddta_sm$treatment <- relevel(as.factor(brooddta_sm$treatment), ref = "control_grp")

gm1 <- lmer(temp ~ ambient + treatment +time1 +  tempIncrease +  (1|colony) + (1|day), data = brooddta_sm, REML = FALSE)

vif.mer(gm1) # doesn't seem too high, when there are no interactions
```


# Conduct more complex analysis and visualization (using a subset of the data)
# GAMM4 -- Generalized Additive Mixed effects Models

<pre>
Here are some notes on specifying models:

The R function used to specify the brood temperature model is as follows: 

gamm4(Brood_Temp ~ s(Ambient_Temp, by = Treatment_Temp_Increase_Interaction, k = 5) + 
s(time_of_day, by = Treatment_Temp_Increase_Interaction, k = 5) +
Treatment_Temp_Increase_Interaction, 
 	random = ~ (1|colony) + (1|day), 
data = brooddta_sm, 
REML = TRUE) 

The formula above shows that a separate smooth was fit for time of day and ambient temperature for each level of the interaction between treatment and temperature direction -- 1) Control group, temp decreasing, 2) Control group, temp increasing, 3) Treatment group, temp decreasing, and 4) treatment group, temp increasing. The final model was fit with Restricted Maximum Likelihood (REML).

The R function used to specify the air temperature model is as follows: 

gamm4(Air_Temp ~ s(Ambient_Temp, by = Treatment_Temp_Increase_Interaction, k = 5) + 
s(time_of_day, by = Treatment_Temp_Increase_Interaction, k = 5) +
Treatment_Temp_Increase_Interaction + cohort, 
random = ~ (1|colony) + (1|day), 
data = airdta_sm, 
REML = TRUE) 

The formula above is the same as the brood model, except that it includes cohort as a predictor and is predicting air temperature inside the colony as the response variable. 
</pre>


```{r, cache = TRUE}
#___________________________________________
# Model relationship between ambient temp and
# brood temp, without transforming the response variable
#___________________________________________
# make a new variable that represents "day" as an integer
# the dayInt variable is unique among cohorts
# cohort 1 starts with day 100, cohort 2 with 200, and 3 with 300
brooddta$dayInt <- as.numeric(brooddta$day) + as.numeric(brooddta$cohort) * 100

numSamps <- 20000


# 20K points is about the max I can do in a reasonable time.
set.seed(17444)
brooddta_sm <- sample_n(brooddta, numSamps, replace = FALSE)
brooddta_sm$treatment <- relevel(as.factor(brooddta_sm$treatment), ref = "control_grp")

# make this into a factor
brooddta_sm$tempIncrease <- as.factor(brooddta_sm$tempIncrease)

# make interaction into a single factor
brooddta_sm$treatTempIncrInt <- interaction(brooddta_sm$treatment, brooddta_sm$tempIncrease)

g00 <- gamm4(temp ~ s(ambient, by = treatment, k = 5) + s(time1, by = treatment, k = 5) + treatment + tempIncrease, random = ~ (1|colony) + (1|dayInt), data = brooddta_sm, REML = FALSE)

# use AIC to compare models
# note: REML must be FALSE to get an accurate AIC for comparison
# note: difference between AIC and AICc is negligible when sample size is large
AICc(g00$mer) 


g02 <- gamm4(temp ~ s(ambient, by = treatTempIncrInt, k = 5) + s(time1, by = treatTempIncrInt, k = 5) + treatTempIncrInt, random = ~ (1|colony) + (1|dayInt), data = brooddta_sm, REML = FALSE)
AICc(g02$mer) # this AICc is lowest


g03 <- gamm4(temp ~ s(ambient, by = treatment, k = 5) + s(time1, by = treatment, k = 5) + treatTempIncrInt, random = ~ (1|colony) + (1|dayInt), data = brooddta_sm, REML = FALSE)
AICc(g03$mer) # this AICc is higher than above

system.time({
  g04 <- gamm4(temp ~ s(ambient, by = treatment, k = 5) + s(time1, by = treatment, k = 5) + treatTempIncrInt + as.factor(cohort), random = ~ (1|colony) + (1|dayInt), data = brooddta_sm, REML = FALSE)
})
AICc(g04$mer) # higher with cohort

# refit with REML=TRUE
g01 <- gamm4(temp ~ s(ambient, by = treatTempIncrInt, k = 5) + s(time1, by = treatTempIncrInt, k = 5) + treatTempIncrInt , random = ~ (1|colony) + (1|dayInt), data = brooddta_sm, REML = TRUE)

```


# Make GAMM figures

```{r}

pdf(file.path(figDir, 'smoothplot_brood.pdf'), width = 15, height = 15)
par(mfrow = c(3,4))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper -- brood
summary(g01$mer)
dev.off()


# plot for html, and for paper
par(mfrow = c(3,4))
plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

{pdf(file.path(figDir, "S10_smoothplot_brood_labels.pdf"), width = 8, height = 8)
par(mfrow = c(3,4))
par(mai= c(0.6,1,0.3,0))

plot.gam(select = 1, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for brood temp, \nControl group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 2, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for brood temp, \nTreatment group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 3, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for brood temp, \nControl group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 4, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for brood temp, \nTreatment group, Temp decreasing, (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 5, xlab = "Time of day (0.5 = noon)", ylab = 'Estimated smooth for brood temp, \nControl group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 6, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for brood temp, \nTreatment group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 7, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for brood temp, \nControl group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 8, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for brood temp, \nTreatment group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

par(mai= c(1.5,1,0.3,0))

plot.gam(select = 9, xlabs = "", ylabs = "Partial residuals\nfor brood temp (°C)", g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE, xaxt = "n")
axis(1, labels = FALSE)
labs <- gsub(x = levels(brooddta_sm$treatTempIncrInt), pattern = "_grp.", replacement = " ")
labs = gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", labs, perl=TRUE) # capitalize
text(cex=1, x=1:4 - 1.5, y=-6.0, labs, xpd=TRUE, srt=45)
text(cex=1, x=3 - 0.75, y=-8.8, "Treatment group\nand temp direction", xpd=TRUE, srt=0)


dev.off()
}



plot(g01$mer)

# residuals are spread a little wide.
mean(residuals(g01$mer) > 2 | residuals(g01$mer) < -2)


# plot actual vs. predicted
brooddta_sm$preds1 <-  predict(g01$mer, type = 'response')
brooddta_sm$preds <-  predict(g01$mer, type = 'response', re.form = NA)
ggplot(brooddta_sm, aes(x = time, y= preds1)) + 
  facet_wrap(~colony + treatment) + 
 geom_line(aes(y = temp), color = 'black') +  # black is actual
 geom_line(alpha = 0.5, color = 'red')  + # red is predicted
  geom_line(alpha = 0.5,aes(y = preds),  color = 'green') # green is predictions without random effects
  

## Overall, model is not obtimal, but it is more interpretable because
# we didn't have to transform the y-variable

# visualize differences
# plot raw data, and prediction, while holding time constant
nd <- brooddta_sm[, c("ambient", "treatment", "time1", "tempIncrease", "treatTempIncrInt")]
nd$time1 = 0.5
nd$ambient <- mean(nd$ambient)

nd <- nd[!(duplicated(nd)), ]
nd

nd2 <- expand.grid(ambient = c(10, 20 , 30), 
                   time1 = seq(0,1, length.out = 30), 
                   treatment = levels(nd$treatment), 
                   treatTempIncrInt = levels(nd$treatTempIncrInt))

nd2$preds1 <-  predict(g01$gam, newdata = nd2, type = 'response', re.form = NA)
nd2$se1 <- predict(g01$gam, newdata = nd2, type = 'response', re.form = NA, se = TRUE)$se


# this shows roughly where the parametric estimates come from
tapply(brooddta_sm$temp, INDEX = list(brooddta_sm$treatTempIncrInt), mean) 
# the intercept is roughly the mean for control group, decreasing temp
summary(g01$gam)$p.table

pdf(file.path(figDir, 'facetedPrededSmooths.pdf'), width = 10, height = 6)
ggplot(nd2, aes(x = time1,  y= preds1, color = as.factor(ambient))) + 
  geom_ribbon(aes(ymin = preds1 - 1.96*se1, ymax = preds1 + 1.96*se1), alpha = 0.4, color = NA) +
  facet_grid(ambient~  treatTempIncrInt , labeller = labeller(.rows = label_both, .cols = label_value)) +
  geom_line(aes(y = preds1)) + 
  #geom_errorbar(aes(ymin = preds1 - 1.96*se1, ymax = preds1 + 1.96*se1), width = 0.05, position = position_dodge(width = 0.02))+ 
  scale_color_viridis(discrete = TRUE, option = "C", end = 0.7, name = "Ambient\nTemp (C)") +
  labs(x = "Time of day (0=midnight, 0.5 = noon)", y="Predicted brood temperature (C)")
dev.off()


brooddta$treatment2 <- mapvalues(brooddta$treatment, from = c("control_grp", "treatment_grp"), 
                                 to = c("No Imidacloprid", "Imidacloprid"))
brooddta$treatment2 <- relevel(as.factor(brooddta$treatment2), ref = "No Imidacloprid")
brooddta$'Temperature Direction' = brooddta$tempIncrease

# save increasing vs. decreasing plots
pdf(file.path(figDir, 'gamSmoothOnIncreasingDecreasing_Brood.pdf'), width = 10, height = 7)
ggplot(sample_n(brooddta, 50000, replace = FALSE), aes(x = ambient,  y= temp, color = interaction( treatment2))) + 
  facet_grid( ~ `Temperature Direction`, labeller = labeller(.rows = label_both, .cols = label_both)) +
  geom_point(aes(y = temp), alpha = 0.05, pch = 16) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) + 
  scale_color_viridis(discrete = TRUE, option = "C", end = 0.7, name = "Treatment") +
  labs(x = "Ambient Temperature (°C)", y="Brood Temperature (°C)")
dev.off()

```

# Output Brood temp tables to HTML for paper
```{r, results='asis'}
# output table to html for paper
s1 <- summary(g01$mer)
# variance for random effects, including smooths
stargazer(as.data.frame(VarCorr(g01$mer), type = 'html'))
write.csv(file = "dataTables/VarRandEffectsBrood.csv", as.data.frame(VarCorr(g01$mer)))
# parametric effects
stargazer(s1$coefficients, type = 'html')

s = summary(g01$gam)
# parametric
stargazer(s$p.table, type = "html")
write.csv(file = "dataTables/parametricBrood.csv", as.data.frame(tidy(s$p.table)))

# smooth terms
stargazer(s$s.table, type = 'html')
write.csv(file = "dataTables/smoothBrood.csv", as.data.frame(tidy(s$s.table)))

```




# GAMM for air temp vs. ambient temp
```{R, cache = TRUE}
#___________________________________________
# Model relationship between ambient temp and
# AIR temp, 
#___________________________________________
# make a new variable that represents "day" as an integer
# the dayInt variable is unique among cohorts
# cohort 1 starts with day 100, cohort 2 with 200, and 3 with 300
airdta$dayInt <- as.numeric(airdta$day) + as.numeric(airdta$cohort) * 100

# 20K points is about the max I can do in a reasonable time.
set.seed(17444)
airdta_sm <- sample_n(airdta, numSamps, replace = FALSE)
airdta_sm$treatment <- relevel(as.factor(airdta_sm$treatment), ref = "control_grp")

# make this into a factor
airdta_sm$tempIncrease <- as.factor(airdta_sm$tempIncrease)

# make interaction into a single factor
airdta_sm$treatTempIncrInt <- interaction(airdta_sm$treatment, airdta_sm$tempIncrease)

g00 <- gamm4(temp ~ s(ambient, by = treatment, k = 5) + s(time1, by = treatment, k = 5) + treatTempIncrInt, random = ~ (1|colony) + (1|dayInt), data = airdta_sm, REML = FALSE)

# use AIC to compare models
# note: REML must be FALSE to get an accurate AIC for comparison
# note: difference between AIC and AICc is negligible when sample size is large
AICc(g00$mer) 


g02 <- gamm4(temp ~ s(ambient, by = treatTempIncrInt, k = 5) + s(time1, by = treatTempIncrInt, k = 5) + treatTempIncrInt, random = ~ (1|colony) + (1|dayInt), data = airdta_sm, REML = FALSE)
AICc(g02$mer) # lowest

g03 <- gamm4(temp ~ s(ambient, by = treatment, k = 5) + s(time1, by = treatment, k = 5) + tempIncrease + treatment, random = ~ (1|colony) + (1|dayInt), data = airdta_sm, REML = FALSE)
AICc(g03$mer) # 

system.time({
g04 <- gamm4(temp ~ s(ambient, by = treatTempIncrInt, k = 5) + s(time1, by = treatTempIncrInt, k = 5) + treatTempIncrInt + as.factor(cohort), random = ~ (1|colony) + (1|dayInt), data = airdta_sm, REML = FALSE)
})
AICc(g04$mer) # lower with cohort


system.time({
  # refit with REML=TRUE
  g01 <- gamm4(temp ~ s(ambient, by = treatTempIncrInt, k = 5) + 
                 s(time1, by = treatTempIncrInt, k = 5) + 
                 treatTempIncrInt + as.factor(cohort), random = ~ (1|colony) + (1|dayInt), 
               data = airdta_sm, REML = TRUE)
})

```


## Make plots for air temp

```{r}

pdf(file.path(figDir, "smoothplot_air.pdf"), width = 15, height = 15)
par(mfrow = c(3,4))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper -- air temp
summary(g01$mer)
dev.off()


# plot for html, and for paper
par(mfrow = c(3,4))
plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

{pdf(file.path(figDir, "S11_smoothplot_air_labels.pdf"), width = 8, height = 8)
par(mfrow = c(3,4))
par(mai= c(0.6,1,0.3,0))

plot.gam(select = 1, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for air temp, \nControl group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 2, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for air temp, \nTreatment group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 3, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for air temp, \nControl group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 4, xlab = "Ambient Temp (°C)", ylab = 'Estimated smooth for air temp, \nTreatment group, Temp decreasing, (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 5, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for air temp, \nControl group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 6, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for air temp, \nTreatment group, Temp decreasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 7, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for air temp, \nControl group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

plot.gam(select = 8, xlab = "Time of day\n(0.5 = noon)", ylab = 'Estimated smooth for air temp, \nTreatment group, Temp increasing (°C)', g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

par(mai= c(1.5,1,0.3,0))

plot.gam(select = 9, xlabs = "", ylabs = "Partial residuals\nfor air temp (°C)", g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE, xaxt = "n")
axis(1, labels = FALSE)
labs <- gsub(x = levels(brooddta_sm$treatTempIncrInt), pattern = "_grp.", replacement = " ")
labs = gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", labs, perl=TRUE) # capitalize
text(cex=1, x=1:4 - 1.5, y=-4.0, labs, xpd=TRUE, srt=45)
text(cex=1, x=3 - 0.75, y=-5.8, "Treatment group\nand temp direction", xpd=TRUE, srt=0)

par(las = 1)
plot.gam(select = 10, xlabs = "Cohort", ylabs = "Partial residuals\nfor air temp (°C)", g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)

dev.off()
}


plot(g01$mer)

# residuals are fine and dandy here!
mean(residuals(g01$mer) > 2 | residuals(g01$mer) < -2)


# plot actual vs. predicted
airdta_sm$preds1 <-  predict(g01$mer, type = 'response')
airdta_sm$preds <-  predict(g01$mer, type = 'response', re.form = NA)
ggplot(airdta_sm, aes(x = time, y= preds1)) + 
  facet_wrap(~colony + treatment) + 
  geom_line(aes(y = temp), color = 'black') +  # black is actual
  geom_line(alpha = 0.5, color = 'red')  + # red is predicted
  geom_line(alpha = 0.5,aes(y = preds),  color = 'green') # green is predictions without random effects


## Overall, this model is easier to fit than brood temp

# visualize differences
# plot raw data, and prediction, while holding time constant
nd <- airdta_sm[, c("ambient", "treatment", "time1", "tempIncrease", "treatTempIncrInt")]
nd$time1 = 0.5
nd$ambient <- mean(nd$ambient)

nd <- nd[!(duplicated(nd)), ]
nd

nd2 <- expand.grid(ambient = c( 10, 20 , 30), 
                   treatTempIncrInt = levels(interaction(levels(nd$treatment), levels(nd$tempIncrease))), 
                   time1 = seq(0,1, length.out = 30), 
                   treatment = levels(nd$treatment), 
                   tempIncrease = levels(nd$tempIncrease), 
                   cohort = unique(airdta_sm$cohort))

nd2$preds1 <-  predict(g01$gam, newdata = nd2, type = 'response', re.form = NA)
nd2$se1 <- predict(g01$gam, newdata = nd2, type = 'response', re.form = NA, se = TRUE)$se

pdf(file.path(figDir, 'facetedPrededSmooths_Air.pdf'), width = 10, height = 6)
ggplot(nd2, aes(x = time1,  y= preds1, color = as.factor(ambient))) + 
  geom_ribbon(aes(ymin = preds1 - 1.96*se1, ymax = preds1 + 1.96*se1), alpha = 0.4, color = NA) +
  facet_grid(ambient~treatTempIncrInt, labeller = labeller(.rows = label_both, .cols = label_value)) +
  geom_line(aes(y = preds1)) + 
  #geom_errorbar(aes(ymin = preds1 - 1.96*se1, ymax = preds1 + 1.96*se1), width = 0.05, position = position_dodge(width = 0.02))+ 
  scale_color_viridis(discrete = TRUE, option = "C", end = 0.7, name = "Ambient\nTemp (°C)") +
  labs(x = "Time of day (0=midnight, 0.5 = noon)", y="Predicted air temperature (C)")
dev.off()



airdta$treatment2 <- mapvalues(airdta$treatment, from = c("control_grp", "treatment_grp"), 
                                 to = c("No imidacloprid", "Imidacloprid"))
airdta$treatment2 <- relevel(as.factor(airdta$treatment2), ref = "No imidacloprid")
airdta$'Temperature direction' = airdta$tempIncrease

# save increasing vs. decreasing plots
pdf(file.path(figDir, 'gamSmoothOnIncreasingDecreasing_Air.pdf'), width = 10, height = 7)
ggplot(sample_n(airdta, 50000, replace = FALSE), aes(x = ambient,  y= temp, color = interaction( treatment2))) + 
  facet_grid( ~ `Temperature direction`, labeller = labeller(.rows = label_both, .cols = label_both)) +
  geom_point(aes(y = temp), alpha = 0.05) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) + 
  scale_color_viridis(discrete = TRUE, option = "C", end = 0.7, name = "Treatment") +
  labs(x = "Ambient temperature (°C)", y="Air temperature\ninside colony (°C)")
dev.off()


# make GAM for both air and brood
data_long$treatment2 <- mapvalues(data_long$treatment, from = c("control_grp", "treatment_grp"), 
                                 to = c("No imidacloprid", "Imidacloprid"))
data_long$treatment2 <- relevel(as.factor(data_long$treatment2), ref = "No imidacloprid")
data_long$'Temperature direction' = data_long$tempIncrease
data_long$'Measurement location' <- mapvalues(data_long$location, from = c("brood", "air"), 
                                to= c("Brood", "Air"))

pdf(file.path(figDir, 'gamSmoothOnIncreasingDecreasing_AirAndBrood.pdf'), width = 10, height = 7)
s7 <- ggplot(sample_n(data_long, 50000, replace = FALSE), aes(x = ambient,  y= temp, color = interaction( treatment2))) + 
  facet_grid(`Measurement location` ~ `Temperature direction`, labeller = labeller(.rows = label_both, .cols = label_both), switch = "y") +
  geom_point(aes(y = temp), alpha = 0.05,pch = 16) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) + 
  scale_color_viridis(discrete = TRUE, option = "C", end = 0.7, name = "Treatment") +
  labs(x = "Ambient temperature (°C)", y="Temperature\ninside colony (°C)") + 
  theme(strip.background = element_blank())
s7
dev.off()

```

# Output HTML tables for air temp

```{r, results='asis'}
s1 <- summary(g01$mer)
# variance for random effects, including smooths
stargazer(as.data.frame(VarCorr(g01$mer), type = 'html'))
write.csv(file = "dataTables/VarRandEffectsAir.csv", as.data.frame(VarCorr(g01$mer)))
# parametric effects
stargazer(s1$coefficients, type = 'html')

s = summary(g01$gam)
# parametric
stargazer(s$p.table, type = "html")
write.csv(file = "dataTables/parametricAir.csv", as.data.frame(tidy(s$p.table)))

# smooth terms
stargazer(s$s.table, type = 'html')
write.csv(file = "dataTables/smoothAir.csv", as.data.frame(tidy(s$s.table)))

```

# combine csv files into one
```{r}
list_of_data <- lapply(dir("dataTables"), FUN = function(x) {
  tmp = read.csv(paste("dataTables", x, sep = "/"), header = FALSE)
                                        tmp$name = x
                                        return(tmp)
                                        })
all_data = do.call(rbind.fill, list_of_data)

write.csv(x = all_data, file = paste("dataTables", "combined_tables_withPermTest.csv", sep = "/"))
```

